---
title: "week 2 assignment"
author: "me"
date: "05/11/2021"
output: html_document
---

## What is mean total number of steps taken per day?

Loading and pre-processing data
```{r}
library(dplyr)

activity <- read.csv("activity.csv")
activity <- subset(activity, !is.na(steps))
activity$date <- as.factor(activity$date)


```

Total number of steps taken per day

```{r}
totalsteps <-unique(activity %>% group_by(date) %>% summarise(date=date, totalsteps=sum(steps)))
totalsteps
```

Histogram of total number of steps per day
```{r}
hist(totalsteps$totalsteps, xlab="Total steps per day", main="Histogram of total steps per day")
oldmean <- mean(totalsteps$totalsteps)
oldmedian <-median(totalsteps$totalsteps)

sum(totalsteps$totalsteps)
```
Mean total no. of steps taken per day = `r oldmean`

Median total no. of steps taken per day = `r oldmedian`

## What is the average daily activity pattern?

```{r}
timeinterval <- unique(activity %>% group_by(interval) %>% summarise(interval=interval, avg_steps=mean(steps)))
plot(avg_steps ~ interval, data=timeinterval, type="l", main="Average daily steps", xlab="day", ylab="Avg steps")

maxinterval <- subset(timeinterval, timeinterval$avg_steps == max(timeinterval$avg_steps))[,1]
```

The 5-min interval with average maximum number of steps across all days is `r maxinterval`.

## Imputing missing values
Reading in the data again to deal with missing values this time.

```{r}
library(lubridate)

activity <- read.csv("activity.csv")
activity$date <- as.Date(activity$date, "%Y-%m-%d")

#adding a new column with the respective weekday
activity$day <- weekdays(activity$date)

#splitting the data by weekdays
daysplit <- split(activity, activity$day)

#writing a function which would impute missing values with the mean for that weekday
impute <- function (y) {
  y[,1][is.na(y[,1])] <- mean(y[,1], na.rm=TRUE)
  y
}
#imputing missing values
daysplitimp <- lapply(daysplit, impute)

#creating the new data frame with no missing values, "newdata"
newdata <- do.call(rbind, daysplitimp)
rownames(newdata) <- NULL

#making new histogram with the total number of steps per day
totalstepsnew <-unique(newdata %>% group_by(date) %>% summarise(date=date, totalsteps=sum(steps)))

hist(totalstepsnew$totalsteps, ylab="Total steps per day", main="Histogram of total steps per day (Imputed)")


newmean <- mean(totalstepsnew$totalsteps); 
newmedian <- median(totalstepsnew$totalsteps)
```

Mean total no. of steps per day (imputed) = `r newmean`
Median total no. of steps per day (imputed) = `r newmedian`

old mean = `r oldmean` new mean = `r newmean`

old median = `r oldmedian` new median = `r newmedian`

## Are there differences in activity patterns between weekdays and weekends?

```{r}
#creating new factor variable for weekend status

newdata <-newdata %>%
  mutate(status= ifelse(day%in% c("Saturday", "Sunday"), "weekend", "weekday"))

newdata$status <- as.factor(newdata$status)

#plotting average no. of steps taken by weekday vs weekend
weekday <- unique(subset(newdata, status=="weekday") %>% 
                    group_by(interval) %>% summarise(interval=interval, avg_steps=mean(steps)))
weekend <- unique(subset(newdata, status=="weekend") %>% 
                    group_by(interval) %>% summarise(interval=interval, avg_steps=mean(steps))) 

par(mfcol=c(2,1))
plot(avg_steps ~ interval, data=weekday, type="l", main="weekday")
plot(avg_steps ~ interval, data=weekend, type="l", main="weekend")

```
